---
title: "Big Problems for Small Networks: Small Network Statistics\\footnote{Contact: \\url{vegayon@usc.edu}. We thank members of our MURI research team, USC's Center for Applied Network Analysis, and Andrew Slaughter for their comments.}"
author:
  - George G. Vega Yon, MS
  - Kayla de la Haye, PhD
date: "North American Social Networks Conference, 2018 \\linebreak[4]November 29, 2018"
output:
  beamer_presentation:
    slide_level: 2 # revealjs::revealjs_presentation
    highlight: espresso
    latex_engine: xelatex
    includes:
      in_header: notation-def.tex
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(smallsize = function(before, options, envir) {
    if (before) {
        "\\footnotesize\n\n"
    } else {
        "\n\\normalsize\n\n"
    }
})
knitr::opts_chunk$set(echo = TRUE, smallsize=TRUE)
```

## Funding Acknowledgement

\begincols
\begincol{.2\linewidth}

\includegraphics[width=1\linewidth]{ARO_logo.png}

\endcol

\begincol{.79\linewidth}
This material is based upon work support by, or in part by, the U.S. Army Research
Laboratory and the U.S. Army Research Office under grant number W911NF-15-1-0577
\endcol

\endcols


\begincols
\begincol{.79\linewidth}

Computation for the work described in this paper was supported by the University
of Southern California’s Center for High-Performance Computing (hpc.usc.edu).
\endcol

\begincol{.2\linewidth}
\includegraphics[width = 1.2\linewidth]{usc.pdf}
\endcol
\endcols


\begin{figure}
\centering
\includegraphics[width = 1\linewidth]{muriteams.png}
\end{figure}

## Context: A tale about social abilities and team performance

Recruited \pause

- 42 mixed gender groups of 3 to 5 participants (unknown)

- Eligibility: (1) 18+ years, (2) Native English speaker \pause

2 hour group session \pause

- Group tasks (2 sets of tasks x 30 minutes each)

- Measure group social networks and individual social system  \pause

Study motivation \pause

- Overall, a very limited set of social has been tested as predictors of social networks

- Very little research on the emergence of networks in teams

## Context  (cont'd)

\begin{figure}
\centering
\includegraphics[width = .9\linewidth]{plot-graph-4-1.pdf}
\end{figure}

\pause

How can we go futher from descriptive statistics? 

<!-- Overall, a very limited set of social has been tested as predictors of social networks -->
<!-- Very little research on the emergence of networks in teams -->

<!-- Aim: Evaluate how a broad range of social intelligence (SI) domains shape team social networks  -->
<!-- Hypotheses:  -->
<!-- teammates with greater SI will be more central in team social networks -->
<!-- teams with more SI will have greater network density, reciprocity and balance -->
<!-- homophily on SI in team social networks -->


## Small networks and Exponential Random Graph Models

When trying to estimate ERGMs in little networks \pause

*   MCMC fails to converge when trying to estimate a block diagonal (structural
    zeros) model,\pause

*   Same happens when trying to estimate an ERGM for a single (little) graph, \pause

*   Even if it converges, the Asymptotic properties of MLEs are no longer valid
    since the sample size is not large enough.

    
## Rethinking the problem

\pause

*   1st Step: Forget about MCMC-MLE estimation, take advantage of small
    sample and use exact statistic for MLEs: \pause
    
    $$
    \Pr\left(\mathbf{Y}=\mathbf{y}|\theta, \mathcal{Y}\right) = \frac{\exp{\theta^{\mbox{T}}\mathbf{g}(\mathbf{y})}}{\kappa\left(\theta, \mathcal{Y}\right)},\quad\mathbf{y}\in\mathcal{Y}
    $$
    
*   This solves the problem of been able to estimate a small ergm. \pause

*   For this we started working on the `lergm` R package
    (available at https://github.com/USCCANA/lergm).

## Example 1

Let's start by trying to estimate an ERGM for a single graph of size 4

```{r lergm1, echo=TRUE}
library(lergm)
set.seed(12)
x <- sna::rgraph(4)
lergm(x ~ edges + balance + mutual)
```


----------

*   Cool, we are able to estimate ERGMs for little networks! (we actually call
    them lergms), but...\pause
    
*   We still have issues regarding asymptotics.\pause

*   We propose to solve this by using a pooled version of the ERGM


## Solution

*   When estimating a block diagonal ERGM we were essentially assuming
    independence across networks.\pause

*   This means that we can actually do the same with exact statistics approach
    to calculate a joint likelihood:\pause
    
    $$
    \Pr\left(\mathbf{Y}=\{\mathbf{y}_{\color{cyan} i}\}|\theta, \left\{\mathcal{Y}_{\color{cyan}i}\right\}\right) = {\color{cyan} \prod_i} \frac{\exp{\theta^{\mbox{T}}\mathbf{g}(\mathbf{y}_{\color{cyan} i})}}{\kappa_{\color{cyan} i}\left(\theta, \mathcal{Y}_{\color{cyan}i}\right)}
    $$
    \pause
    
*   By estimating a pooled version of the ERGM we can recover the asymptotics 
    of MLEs.\pause
    
*   We implemented this in the `lergm` package
    
## Example 2

Suppose that we have 3 little graphs of sizes 4, 5, and 5:

```{r lergm2, echo=TRUE}
library(lergm)
set.seed(12)
x1 <- sna::rgraph(4)
x2 <- sna::rgraph(5)
x3 <- sna::rgraph(5)

lergm(list(x1, x2, x3) ~ edges + balance + mutual)
```

<!-- ## Convergence diagnostics -->

<!-- ```{r diagnostics2} -->
<!-- plot(ans) -->
<!-- ``` -->

## Simulation study

### Scenario A

1.  Draw parameters for _edges_ and _mutual_ from a uniform(-3, 3).

2.  Using those parameters, sampled $n\sim\mbox{Poisson}(30)$ networks of size 4

3.  Estimated the pooled ERGMs using both the MLE and the boostrap version.

\pause

### Scenario B

1.  Idem.

2.  Using those parameters, sampled $n_1\sim\mbox{Poisson}(15), n_2\sim\mbox{Poisson}(15)$
    networks of size 3 and 4 respectively.

3.  Idem.

## Simulation study: Scenario A

### Empirical Bias

\begin{figure}
\centering
\includegraphics[width=.8\linewidth]{bias-01-fixed-sizes-4.pdf}
\end{figure}

## Simulation study: Scenario A

### Power

\begin{figure}
\centering
\includegraphics[width=.8\linewidth]{power-01-fixed-sizes-4.pdf}
\end{figure}


## Simulation study: Scenario B

### Empirical Bias

\begin{figure}
\centering
\includegraphics[width=.8\linewidth]{bias-02-various-sizes-3-4.pdf}
\end{figure}

## Simulation study: Scenario B

### Power

\begin{figure}
\centering
\includegraphics[width=.8\linewidth]{power-02-various-sizes-3-4.pdf}
\end{figure}

## Other approaches

\begin{figure}
\centering
\includegraphics[height=.9\textheight]{similr.png}
\end{figure}

## Discussion

*   First set of results from the simulation study are encouraging\pause

*   Need to conduct more simulations using _nodal_ attributes and networks of size 5
    (right now having problems when building the DGP).\pause

*   Small structures imply a smaller pool of parameters (which is OK), but 
    can be more useful when including nodal attributes.\pause
    
*   We have to work on parameter interpretation. Right now model statistics are
    not centered as these are in _regular_ ERGMs (perhaps using the means?).\pause
    
*   When estimating the pooled version, we are essentially hand-waving the fact
    that parameter estimates implicitly encode size of the graph, i.e.
    
    > Does a the estimate of `edge = 0.1` has the same meaning for a network of
    size 3 to a size 5?


##


### Thank you!

\maketitle
