---
title: "Big Problems for Small Networks: Small Network Statistics"
author:
  - George G. Vega Yon, MS
  - Kayla de la Haye, PhD
date: "North American Social Networks Conference, 2018 \\linebreak[4]November 29, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(smallsize = function(before, options, envir) {
    if (before) {
        "\\footnotesize\n\n"
    } else {
        "\n\\normalsize\n\n"
    }
})
knitr::opts_chunk$set(echo = TRUE, smallsize=TRUE)
```

## Context

*   We have 4 teams with 3 to 5 individuals.

*   We have information about the individuals (gender, age, GPA, etc.)

*   Individuals were randomly assigned to teams with the criterion that there was at least one male and one female per group

*   Each team was asked to perform a series of 12 group tasks (using 2 versions of MIT's Test for Collective Intelligence); and each teammate individually reported on (A) a comprehensive measure of personality, and emotional/social intelligence, and (B) their perceptions of the team social network  (multiplex tie)s

We would like to be able to estimate ERGM models here.

## The problem with Small networks

*   MCMC breaks (no convergence) when trying to estimate a block diagonal model,

*   Same happens when trying to estimate an ERGM for a single (little) graph.

*   Even if it converges, the Asymptotic properties of MLEs are no longer valid
    since the sample size is not large enough.

    
## Rethinking the problem

*   1st Step: Forget about MCMC-MLE estimation, take advantage of small
    sample and use exact statistic for MLEs.
    
*   This solves the problem of been able to estimate a small ergm.

*   For this we started working on the `lergm` R package (available at https://github.com/USCCANA/lergm).

## Example 1

Let's start by trying to estimate an ERGM for a single graph of size 4

```{r lergm1, echo=TRUE}
library(lergm)
set.seed(12)
x <- sna::rgraph(4)
lergm(x~edges + balance)
```


----------

*   Cool, we are able to estimate ERGMs for little networks! (we actually call
    them lERGMs), but...
    
*   We still have issues regarding asymptotics.

*   We propose to solve this by using a pulled version of the ERGM


## Solution

*   Remember that we were trying to estimate a block diagonal ERGM?
    
    We were essentially assuming independence across teams

*   This means that we can actually do the same with exact statistics
    
    $$
    \Pr\left(\mathbf{Y}={\color{cyan} \{}\mathbf{y}_{\color{cyan} i}{\color{cyan} \}}|\theta, \mathcal{Y}\right) = {\color{cyan} \prod_i} \frac{\exp{\theta^{\mbox{T}}\mathbf{g}(\mathbf{y}_{\color{cyan} i})}}{\kappa_{\color{cyan} i}\left(\theta, \mathcal{Y}\right)}
    $$
    
*   By estimating a pulled version of the ERGM (which is equivalent to block diagonal),
    we can recover the asymptotics of MLEs.
    
*   We implemented this in the `lergm` package
    
## Example 2

Suppose that we have 3 little graphs of sizes 4, 5, and 5:

```{r lergm2, echo=TRUE}
library(lergm)
set.seed(12)
x1 <- sna::rgraph(4)
x2 <- sna::rgraph(5)
x3 <- sna::rgraph(5)

ans <- lergm(list(x1, x2, x3) ~ edges + balance)
```

One of the current problems is the inestability of the likelihood function

## Convergence diagnostics

```{r diagnostics2}
plot(ans)
```

---

Multiple runs may give different results

```{r inestability1}
set.seed(1)
lergm(list(x1, x2, x3) ~ edges + balance, stats = ans$model$stats)
lergm(list(x1, x2, x3) ~ edges + balance, stats = ans$model$stats)
```


## Discussion

*   This is no panacea: The external validity lERGMs inference is yet more complicated than ERGMs.

*   Small structures yield a smaller pool of parameters... unless we use nodes'
    covariates (otherwise the model is not ~~very~~ informative)
    
*   When estimating the pooled version, we are essentially hand-waving the fact
    that parameter estimates implicitly encode size of the graph, i.e.
    
    > Does a the estimate of `edge = 0.1` has the same meaning for a network of
    size 3 to a size 5?

*   We could actually go further and think about Separable Exponential Random
    Graph Models (aka TERGMs)... this could be an interesting approach.

##

### Thank you!

\maketitle

Research Supported by: